Reproducible methods for modeling combined public transport and cycling trips and associated benefits: evidence from the biclaR tool

### 2024-05-08 paper/PaperCEUS/PaperCEUS_rev1.Rmd
review FM + diff

This paper estimates the potential for combining cycling and PT to substitute car trips in the LMA.
After presenting the methods used, we assess its socio-environmental impacts using open data and open-source tools.

@@

The Lisbon metropolitan area (LMA), in Portugal, is the case study, comprising 18 municipalities, with a total area of 3,015 km² and a population of 2.8 million residents
The LMA is the most populous and economically active region in Portugal, with the city of Lisbon as the central node of the LMA, with a high concentration of jobs and services, and a high demand for mobility (see Figure \ref{fig:lma}).

@@

```{r lma, fig.cap="\\label{lma}Population and daily trips (origins) at each Municipality of the LMA, according to census and the travel survey.", fig.subcap=c("Population distribution in the LMA","Trips/day in the LMA"), fig.ncol=2, message=FALSE, warning=FALSE, out.width="50%"}
MUNICIPIOSgeo = readRDS("load/MUNICIPIOSgeo.Rds")
MUNICIPIOSgeo_mapa = readRDS("load/MUNICIPIOSgeo_mapa.Rds")
MUNICIPIOS_dtcc = readRDS("load/MUNICIPIOS_dtcc.Rds")
POPMunicipios = readRDS("load/POPMunicipios.Rds")

@@

```{r mapzones, fig.cap="\\label{mapzones}Potential Bicycle + PT daily trips for the 10% target and direct routing, by municipality. The percentages are shown for municipalities with values above 4%.", warning=FALSE, fig.subcap=c("Potential new Bicycle + PT trips","Proportion of total potential modal shift in the LMA"), fig.ncol=2, message=FALSE, out.width="50%"}
# plot 1
ggplot(data = CAR_PT_TRANSF_mun_or_geo_18_extrap) +
  geom_sf(aes(fill = Bike10)) +
  viridis::scale_fill_viridis(option = "G",
                              direction = -1,
                              name = "Potential new \nBike + PT\ntrips/day",
                              na.value = "white",
                              labels = scales::unit_format(unit = "k", scale = 1e-3)) +

@@

# # plot 2 - the total % of potential bike + TP trips
# ggplot(data = CAR_PT_TRANSF_mun_or_geo_18) +
#   geom_sf(aes(fill = TP_10per)) +
#   viridis::scale_fill_viridis(option = "G", direction = -1, name = "Bike + PT trips %", na.value = "white") +
#   theme_void()
# 
# # plot 3 - the added % of potential bike + TP trips - doesn't make sense to show
# ggplot(data = CAR_PT_TRANSF_mun_or_geo_18) +
#   geom_sf(aes(fill = TP_10per- TPper)) + #???
#   viridis::scale_fill_viridis(option = "G", direction = -1, name = "new Bike + PT trips %", na.value = "white") +
#   theme_void()

@@

# Generalizability and limitations

@@

# Conclusion

@@

The findings indicate that cycling combined with public transport has the potential to account for 10% of total trips. Transitioning to cycling for the first and last segments could lead to a decrease in annual CO~2~eq emissions by a substantial margin, ranging from 3,000 to 7,500 tons per day. Moreover, for the second trip segment, the shift from car to public transport contributes to avoiding up to 20,500 tons of CO~2~eq emissions annually. These changes are estimated to yield socio-environmental benefits totaling €125 million to €325 million over a decade.
The quantification of such benefits can support policy-makers in prioritizing interventions to reduce the reliance on private motorized modes of transportation.

Additionally, the results suggest that the potential for cycling-PT intermodality is spatially unevenly distributed across the LMA, with the highest potential in the municipalities with the highest number of trips and PT interfaces of PT modes connecting to Lisbon and where it is possible to carry bicycles on board, such as Lisbon, Sintra, Amadora, and Almada.

We opted to detail all the steps of the methodology, from data requirements to the final results, to ensure transparency and generalization of the approach.
We acknowledge the limitations of the used methods and propose avenues for future research.

>>>>

This paper estimates the potential for combining cycling and PT to substitute car trips in the LMA.
After presenting the [Methods] used in Chapter 2, we assess its socio-environmental impacts using open data and open-source tools, presenting the [Results and Discussion] in Chapter 3, followed by Chapter 4 where we discuss the [Generalizability and Limitations] of the methods.
We [Conclude](#Conclusions) the paper in Chapter 5 with the main findings and policy implications.

@@

The Lisbon metropolitan area (LMA), in Portugal, is the case study, comprising 18 municipalities, with a total area of 3,015 km² and a population of 2.8 million residents, according to the 2021 census [@INEcensus].
The LMA is the most populous and economically active region in Portugal, with the city of Lisbon as the central node of the LMA, with a high concentration of jobs and services, and a high demand for mobility (see Figure \ref{fig:lma}).

@@

```{r lma, fig.cap="\\label{lma}Population and daily trips (origins) at each Municipality of the LMA, according to the 2021 census and the 2017 travel survey.", fig.subcap=c("Population distribution in the LMA","Trips/day in the LMA"), fig.ncol=2, message=FALSE, warning=FALSE, out.width="50%"}
MUNICIPIOSgeo = readRDS("load/MUNICIPIOSgeo.Rds")
MUNICIPIOSgeo_mapa = readRDS("load/MUNICIPIOSgeo_mapa.Rds")
MUNICIPIOS_dtcc = readRDS("load/MUNICIPIOS_dtcc.Rds")
POPMunicipios = readRDS("load/POPMunicipios.Rds")

@@

```{r mapzones, fig.cap="\\label{mapzones}Potential Bicycle + PT daily trips for the 10% target and direct routing, by municipality (origins). The percentages are shown for municipalities with values above 4%.", warning=FALSE, fig.subcap=c("Potential new Bicycle + PT trips","Proportion of total potential modal shift in the LMA"), fig.ncol=2, message=FALSE, out.width="50%"}
# plot 1
ggplot(data = CAR_PT_TRANSF_mun_or_geo_18_extrap) +
  geom_sf(aes(fill = Bike10)) +
  viridis::scale_fill_viridis(option = "G",
                              direction = -1,
                              name = "Potential new \nBike + PT\ntrips/day",
                              na.value = "white",
                              labels = scales::unit_format(unit = "k", scale = 1e-3)) +

@@

# # plot 2 - the total % of potential bike + TP trips
# ggplot(data = CAR_PT_TRANSF_mun_or_geo_18) +
#   geom_sf(aes(fill = TP_10per)) +
#   viridis::scale_fill_viridis(option = "G", direction = -1, name = "Bike + PT trips %", na.value = "white") +
#   theme_void()
# 
# # plot 3 - the added % of potential PT
# ggplot(data = CAR_PT_TRANSF_mun_or_geo_18) +
#   geom_sf(aes(fill = TP_10per- TPper)) + #???
#   viridis::scale_fill_viridis(option = "G", direction = -1, name = "new Bike + PT trips %", na.value = "white") +
#   theme_void()

@@

# Generalizability and Limitations

@@

# Conclusion {#Conclusions}

@@

The findings indicate that cycling combined with public transport has the potential to account for 10% of total trips. Transitioning to cycling for the first and last segments could lead to a decrease in annual CO~2~eq emissions by a substantial margin, ranging from 3,000 to 7,500 tons per day, depending on the cycling target and routing profile. Moreover, for the second trip segment, the shift from car to public transport contributes to avoiding up to 20,500 tons of CO~2~eq emissions annually. These changes are estimated to yield socio-environmental benefits totaling €125 million to €325 million over a decade.
The quantification of such benefits can support policy-makers in prioritizing interventions to reduce the reliance on private motorized modes of transportation.

Additionally, the results suggest that the potential for cycling-PT intermodality is spatially unevenly distributed across the LMA. The highest potential occurs in the municipalities with the highest number of trips and PT interfaces with connecting modes to Lisbon that allow for carrying bicycles on board, such as Lisbon, Sintra, Amadora, and Almada.

We opted to detail all the steps of the methodology, from data requirements to the final results, to ensure transparency and generalization of the approach, for instance making sensitivity analysis to the bike trip distances and the PT number of transfers.
We acknowledge the limitations of the used methods and propose avenues for future research.

### 2024-05-08 paper/PaperCEUS/PaperCEUS_rev1.tex
review FM + diff

This paper estimates the potential for combining cycling and PT to
substitute car trips in the LMA. After presenting the methods used, we
assess its socio-environmental impacts using open data and open-source
tools.

@@

The Lisbon metropolitan area (LMA), in Portugal, is the case study,
comprising 18 municipalities, with a total area of 3,015 km² and a
population of 2.8 million residents The LMA is the most populous and
economically active region in Portugal, with the city of Lisbon as the
central node of the LMA, with a high concentration of jobs and services,
and a high demand for mobility (see Figure \ref{fig:lma}).

@@

\begin{figure}
\subfloat[Population distribution in the LMA\label{fig:lma-1}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/lma-1} }\subfloat[Trips/day in the LMA\label{fig:lma-2}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/lma-2} }\caption{\label{lma}Population and daily trips (origins) at each Municipality of the LMA, according to census and the travel survey.}\label{fig:lma}
\end{figure}

@@

\begin{figure}
\subfloat[Potential new Bicycle + PT trips\label{fig:mapzones-1}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/mapzones-1} }\subfloat[Proportion of total potential modal shift in the LMA\label{fig:mapzones-2}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/mapzones-2} }\caption{\label{mapzones}Potential Bicycle + PT daily trips for the 10\% target and direct routing, by municipality. The percentages are shown for municipalities with values above 4\%.}\label{fig:mapzones}
\end{figure}

@@

\section{Generalizability and
limitations}\label{generalizability-and-limitations}

@@

\section{Conclusion}\label{conclusion}

@@

The findings indicate that cycling combined with public transport has
the potential to account for 10\% of total trips. Transitioning to
cycling for the first and last segments could lead to a decrease in
annual CO\textsubscript{2}eq emissions by a substantial margin, ranging
from 3,000 to 7,500 tons per day. Moreover, for the second trip segment,
the shift from car to public transport contributes to avoiding up to
20,500 tons of CO\textsubscript{2}eq emissions annually. These changes
are estimated to yield socio-environmental benefits totaling €125
million to €325 million over a decade. The quantification of such
benefits can support policy-makers in prioritizing interventions to
reduce the reliance on private motorized modes of transportation.

Additionally, the results suggest that the potential for cycling-PT
intermodality is spatially unevenly distributed across the LMA, with the
highest potential in the municipalities with the highest number of trips
and PT interfaces of PT modes connecting to Lisbon and where it is
possible to carry bicycles on board, such as Lisbon, Sintra, Amadora,
and Almada.

We opted to detail all the steps of the methodology, from data
requirements to the final results, to ensure transparency and
generalization of the approach. We acknowledge the limitations of the
used methods and propose avenues for future research.

>>>>

This paper estimates the potential for combining cycling and PT to
substitute car trips in the LMA. After presenting the
\hyperref[methods]{Methods} used in Chapter 2, we assess its
socio-environmental impacts using open data and open-source tools,
presenting the \hyperref[results-and-discussion]{Results and Discussion}
in Chapter 3, followed by Chapter 4 where we discuss the
\hyperref[generalizability-and-limitations]{Generalizability and
Limitations} of the methods. We \hyperref[Conclusions]{Conclude} the
paper in Chapter 5 with the main findings and policy implications.

@@

The Lisbon metropolitan area (LMA), in Portugal, is the case study,
comprising 18 municipalities, with a total area of 3,015 km² and a
population of 2.8 million residents, according to the 2021 census
\citep{INEcensus}. The LMA is the most populous and economically active
region in Portugal, with the city of Lisbon as the central node of the
LMA, with a high concentration of jobs and services, and a high demand
for mobility (see Figure \ref{fig:lma}).

@@

\begin{figure}
\subfloat[Population distribution in the LMA\label{fig:lma-1}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/lma-1} }\subfloat[Trips/day in the LMA\label{fig:lma-2}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/lma-2} }\caption{\label{lma}Population and daily trips (origins) at each Municipality of the LMA, according to the 2021 census and the 2017 travel survey.}\label{fig:lma}
\end{figure}

@@

\begin{figure}
\subfloat[Potential new Bicycle + PT trips\label{fig:mapzones-1}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/mapzones-1} }\subfloat[Proportion of total potential modal shift in the LMA\label{fig:mapzones-2}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_rev1_files/figure-latex/mapzones-2} }\caption{\label{mapzones}Potential Bicycle + PT daily trips for the 10\% target and direct routing, by municipality (origins). The percentages are shown for municipalities with values above 4\%.}\label{fig:mapzones}
\end{figure}

@@

\section{Generalizability and
Limitations}\label{generalizability-and-limitations}

@@

\section{Conclusion}\label{Conclusions}

@@

The findings indicate that cycling combined with public transport has
the potential to account for 10\% of total trips. Transitioning to
cycling for the first and last segments could lead to a decrease in
annual CO\textsubscript{2}eq emissions by a substantial margin, ranging
from 3,000 to 7,500 tons per day, depending on the cycling target and
routing profile. Moreover, for the second trip segment, the shift from
car to public transport contributes to avoiding up to 20,500 tons of
CO\textsubscript{2}eq emissions annually. These changes are estimated to
yield socio-environmental benefits totaling €125 million to €325 million
over a decade. The quantification of such benefits can support
policy-makers in prioritizing interventions to reduce the reliance on
private motorized modes of transportation.

Additionally, the results suggest that the potential for cycling-PT
intermodality is spatially unevenly distributed across the LMA. The
highest potential occurs in the municipalities with the highest number
of trips and PT interfaces with connecting modes to Lisbon that allow
for carrying bicycles on board, such as Lisbon, Sintra, Amadora, and
Almada.

We opted to detail all the steps of the methodology, from data
requirements to the final results, to ensure transparency and
generalization of the approach, for instance making sensitivity analysis
to the bike trip distances and the PT number of transfers. We
acknowledge the limitations of the used methods and propose avenues for
future research.
